---
title: "Data_analysis_in_R_PhD_class_JanosSalamon"
author: "Janos Salamon"
date: "June 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive summary 

The present study investigates the connection between burnout, work satisfaction and
different kind of work motivation factors. Multidimensional Work Motivation Scale (MWMS; Gagné et al., 2015), 
Satisfaction with Work Scale (SWWS; Fouquereau & Rioux, 2002) and 
Shirom-Melamed Burnout Questionnaire (SMBQ; Shirom & Melamed, 2006) were used in the study. 
The study relies on a sample of 721 Hungarian working adults (532 female, 73.8%), who were aged 
between 19 and 71 (M =  37.4; SD = 11.6). These participants reported their 
highest level of education as primary (1.4%), secondary (34.1%) and higher (64.5%). 
The majority of the sample works more than 15 years ago (43.3%), 30% works between 4 and 15 years ago, while 
24.7% works less than 4 years ago. There are participants (7.2%) who are self-employed, 
the majority works at SMEs (55.1%), and 37.7% works at large enterprises. 
The aim of the study was first to assess each constructs factor structure 
using confirmatory factor analysis (CFA), then investigate the connection 
between these factors, using structural equation modeling (SEM). 
As fit indeces did not show acceptable results for the factors of Multidimensional Work Motivation Scale (MWMS), 
it was necessary to continue analysis and explore factor structure by using Principal Components Analysis (PCA).
PCA of MWMS suggested a 5-factor model, while there was important to drop out 2 items for a better factor structure and acceptable fit indices. Fit indices of the final model are satisfactory.

# Preparation
Loading packages
```{r}
library(tidyverse)
library(lavaan)
library(haven)
```

```{r, include=FALSE}
# set working direcotry and import data from SPSS (.sav) file
setwd("C:/Users/salam/Desktop/Tanulas/R kurzus")
work_data <- read_sav("munka_master1.sav")
```


## 1. Confirmatory Factor Analysis 
Confirmatory factor analysis using the lavaan package

# model for burnout
```{r}
burnout <- '
  Physical_B =~ burn_1 + burn_3 + burn_5 + burn_8 + burn_10 + burn_12
  Cognitive_B =~ burn_2 + burn_6 + burn_9 + burn_11 + burn_13
  Emotional_B =~ burn_4 + burn_7 + burn_14'
fit_burnout <- cfa(burnout, data = work_data, estimator = 'MLR')
summary(fit_burnout, fit.measures = TRUE, standardized = TRUE)
```
Fit indices highlighted that the hypothesized representation of the constructs 
is satisfactory (CFI = .924, TLI = .907, RMSEA = .082 [90% CI = .075-.090]).


# model for work motivation
```{r}
work_motiv <- '
  Amotiv =~ mwms_4 + mwms_10 + mwms_16
  Ext_Mat =~ mwms_6 + mwms_12 + mwms_18
  Ext_Soc =~ mwms_1 + mwms_7 + mwms_13
  Introj =~ mwms_2 + mwms_8 + mwms_14 + mwms_19
  Ident =~ mwms_5 + mwms_11 + mwms_17
  Intrin =~ mwms_3 + mwms_9 + mwms_15'
fit_work_motiv <- cfa(work_motiv, data = work_data, estimator = 'MLR')
summary(fit_work_motiv, fit.measures = TRUE, standardized = TRUE)
```
Fit indices show that the hypothesized representation of the constructs 
is unsatisfactory (CFI = .907, TLI = .884, RMSEA = .069 [90% CI = .064-.074]).


# model for satisfaction
```{r}
work_satisf <- ' work_satisf =~ swws_1 + swws_2 + swws_3 + swws_4 + swws_5
  swws_2 ~~ swws_5'
fit_work_satisf <- cfa(work_satisf, data = work_data, estimator = 'MLR')
summary(fit_work_satisf, fit.measures = TRUE, standardized = TRUE, modindices = TRUE)
```
Fit indices show that the hypothesized representation of the constructs 
is good (CFI = .998, TLI = .996, RMSEA = .029 [90% CI = .000-.065]), 
taking into account the simlicity of the model, and the range of the confidence inteval.

Modification Indices suggested error covariances between the following items: swws_2 and swws_5,
thus these are corrected in the final model.


## SEM model
```{r}
my_model <- ' 
  # measurement model
    burnout =~ Physical_B + Emotional_B + Cognitive_B
                  Physical_B =~ burn_1 + burn_2 + burn_3 + burn_4 + burn_5 + burn_6
                  Emotional_B =~ burn_7 + burn_8 + burn_9 + burn_10 + burn_11
                  Cognitive_B =~ burn_12 + burn_13 + burn_14

    Amotiv =~ mwms_4 + mwms_10 + mwms_16
    Ext_Mat =~ mwms_6 + mwms_12 + mwms_18
    Ext_Soc =~ mwms_1 + mwms_7 + mwms_13
    Introj =~ mwms_2 + mwms_8 + mwms_14 + mwms_19
    Ident =~ mwms_5 + mwms_11 + mwms_17
    Intrin =~ mwms_3 + mwms_9 + mwms_15
    
    work_satisf =~ swws_1 + swws_2 + swws_3 + swws_4 + swws_5
  
# regressions
    burnout ~ Amotiv + Ext_Mat + Ext_Soc + Introj + Ident + Intrin
    work_satisf ~ Amotiv + Ext_Mat + Ext_Soc + Introj + Ident + Intrin
    burnout ~~ work_satisf'
```

# model fit and summary of model fit
```{r}
my_model_fit <- cfa(my_model, data = work_data, estimator = 'MLR')
```
Output shows warning message: "some estimated lv variances are negative", 
which was expected based on previous results.

```{r}
summary(my_model_fit, fit.measures=TRUE, standardized = TRUE)
```
Fit indices highlighted that the hypothesized representation of the constructs 
is unsatisfactory (CFI = .858, TLI = .843, RMSEA = .065 [90% CI = .062-.068]).
These results were expected based on the fit indices 
of some elements of the model (showed in previous results).


# Plotting the models with semPlot
```{r}
library(semPlot)
```

# Seeing the clean model structure
```{r}
semPaths(my_model_fit)
```

# Getting standardized estimates
```{r}
semPaths(my_model_fit, "std")
```

The model shows the previously indicated negative correlations between some variables.
Based on the unsatisfactory fit indices of previously hypothesized representation of the constructs 
it is suggested to continue the analysis with exploratory factor analysis. 


## 2. Dimension reduction (Principal Components Analysis)

PCA for exploring the factor sturcture of motivation, using the Multidimensional Work Motivation Scale, 
because fit indices in CFA were unsatisfactory in the original factor structure.

# Loading packages
```{r}
library(corpcor) 
library(GPArotation)
library(psych)
library(devtools)
library(DSUR.noof)
```

```{r}
motiv <- 
  work_data %>%
  as_tibble() %>%
  select(mwms_1:mwms_19)
```


# Preliminary analysis for EFA
```{r}
cortest.bartlett(motiv, n = 721)
```
Bartlett's test is significant, x2(171) = 6,720, p < .001), so factor analysis is appropriate.

```{r}
kmo(motiv)
```
KMO statistics show that overall KMO is 0.86, which means: 
"The KMO test yields a degree of common variance meritorious."
KMO values for Individual variables also acceptable (all are > .5). 

```{r}
det(cor(motiv))
```
Determinant of the correlation matrix (8.046915e-05) is greater than 0.00001, 
thus determinant does not seem problematic either.

# Finding the optimal number of factors
```{r}
pc1 <- principal(motiv, nfactors = 19, rotate = "none")
pc1

plot(pc1$values, type = "b")
```

Scree plot shows an inflexion after 6 factors, while 
eigenvalues, following Jolliffe's criterion also supports the six-component solution.
Using promax rotation, because it is theoreticaly grounded that 
motivational factors might correlate.

```{r}
pc2 <- principal(motiv, nfactors = 6, rotate = "promax")
pc2
print.psych(pc2, cut = 0.3, sort = TRUE)
```
6-factor model's fit shows upon off diagonal values = 0.97, which means a good fit.
Factor structure does not clear, constructs are also mixed in a not meaningful way.

# Trying to check a 5-factor model, based on Keiser's criterion.
```{r}
pc3 <- principal(motiv, nfactors = 5, rotate = "promax")
pc3
print.psych(pc3, cut = 0.3, sort = TRUE)
```
5-factor model's fit also shows upon off diagonal values = 0.97, which means a good fit.
Pattern matrix shows high cross-loadings of item 14.

# Deleting item 14, then re-run analysis.
```{r}
library(tidyverse)

motiv2 <- 
  work_data %>%
  as_tibble() %>%
  select(mwms_1, mwms_2, mwms_3, mwms_4, mwms_5, mwms_6, mwms_7,
         mwms_8, mwms_9, mwms_10, mwms_11, mwms_12, mwms_13, mwms_15,
         mwms_16, mwms_17, mwms_18, mwms_19)


pc4 <- principal(motiv2, nfactors = 5, rotate = "promax")
pc4
print.psych(pc4, cut = 0.3, sort = TRUE)
```
Pattern matrix shows high cross-loadings of item 13.

# Deleting item 13, then re-run analysis.
```{r}
motiv3 <- 
  work_data %>%
  as_tibble() %>%
  select(mwms_1, mwms_2, mwms_3, mwms_4, mwms_5, mwms_6, mwms_7,
         mwms_8, mwms_9, mwms_10, mwms_11, mwms_12, mwms_15,
         mwms_16, mwms_17, mwms_18, mwms_19)

pc5 <- principal(motiv3, nfactors = 5, rotate = "promax")
pc5
print.psych(pc5, cut = 0.3, sort = TRUE)
```
Pattern matrix shows very clear loadings. Factors are theoretically appropriate.


# Variables are loading to the following factors. Each show a meaningful factor:  
  factor 1: Intinsic motivation (Intrin): mwms_3 + mwms_9 + mwms_15
  
  factor 2: Autonomous extrinsic (Aut_Ext): mwms_5 + mwms_11 + mwms_17 + mwms_19
  
  factor 3: Amotivation (Amotiv): mwms_4 + mwms_10 + mwms_16
  
  fcator 4: Controlled extrinsic (Cont_Ext): mwms_1 + mwms_2 + mwms_7 + mwms_8
  
  factor 5: External material (Ext_Mat): mwms_6 + mwms_12 + mwms_18



## 3. CFA model for new work motivation factors
```{r}
library(lavaan)

work_motiv <- '
Intrin =~ mwms_3 + mwms_9 + mwms_15
Aut_Ext =~ mwms_5 + mwms_11 + mwms_17 + mwms_19
Amotiv =~ mwms_4 + mwms_10 + mwms_16
Cont_Ext =~ mwms_1 + mwms_2 + mwms_7 + mwms_8
Ext_Mat =~ mwms_6 + mwms_12 + mwms_18'

fit_work_motiv <- cfa(work_motiv, data = work_data, estimator = 'MLR')
summary(fit_work_motiv, fit.measures = TRUE, standardized = TRUE)
```

Fit indices of the hypothesized representation of the constructs 
are satisfactory now (CFI = .941, TLI = .927, RMSEA = .057 [90% CI = .051-.063]).



## 4. Normality
Identifing violations of normality

# Loading packages
```{r}
library(dplyr)
library(MVN)
```

# variable calculation for work motivation factors
```{r}
motiv_variables <-
  work_data %>%
  mutate(Intrin_avg = mean(mwms_3, mwms_9, mwms_15, trim = 0, na.rm=TRUE),
         Aut_Ext_avg = mean(mwms_5, mwms_11, mwms_17, trim = 0, mwms_19, na.rm=TRUE),
         Amotiv_avg = mean(mwms_4, mwms_10, mwms_16, trim = 0, na.rm=TRUE),
         Cont_Ext_avg = mean(mwms_1, mwms_2, mwms_7, trim = 0, mwms_8, na.rm=TRUE),
         Ext_Mat_avg = mean(mwms_6, mwms_12, mwms_18, trim = 0, na.rm=TRUE))
```

# assessment of univariate and multivariate normality with Mardia's test - work motivation
```{r}
avg_only_motiv <- motiv_variables %>%
  select(mwms_1, mwms_2, mwms_3, mwms_4, mwms_5, mwms_6, mwms_7, mwms_8, mwms_9, 
         mwms_10, mwms_11, mwms_12, mwms_15, mwms_16, mwms_17, mwms_18, mwms_19)

mvn(avg_only_motiv)
```

# Filtering missing values from burnout data
```{r}
work_data2 <- work_data %>%
  filter(!is.na(burn_1), 
         !is.na(burn_2),
         !is.na(burn_3),
         !is.na(burn_4),
         !is.na(burn_5),
         !is.na(burn_6),
         !is.na(burn_7),
         !is.na(burn_8),
         !is.na(burn_9),
         !is.na(burn_10),
         !is.na(burn_11),
         !is.na(burn_12),
         !is.na(burn_13),
         !is.na(burn_14))
```

# variable calculation for burnout factors
```{r}
burnout_variables <-
  work_data2 %>%
  mutate(Physical_B_avg = mean(burn_1, burn_3, burn_5, burn_8, burn_10, burn_12, trim = 0, na.rm=TRUE),
         Cognitive_B_avg = mean(burn_2, burn_6, burn_9, burn_11, burn_13, trim = 0, na.rm=TRUE),
         Emotional_B_avg = mean(burn_4, burn_7, burn_14, trim = 0, na.rm=TRUE))
```

# assessment of univariate and multivariate normality with Mardia's test - burnout
```{r}
avg_only_burn <- burnout_variables %>%
  select(burn_1, burn_2, burn_3, burn_4, burn_5, burn_6, burn_7, burn_8, burn_9, burn_10, 
         burn_11, burn_12, burn_13, burn_14)

mvn(avg_only_burn)
```

# Filtering missing values from satisfaction data
```{r}
work_data3 <- work_data %>%
  filter(!is.na(swws_1), 
         !is.na(swws_2),
         !is.na(swws_3),
         !is.na(swws_4),
         !is.na(swws_5))
```

# variable calculation for work satisfaction factor
```{r}
sat_variables <-
  work_data3 %>%
  mutate(satis_avg = mean(swws_1, swws_2, swws_3, swws_4, swws_5, trim = 0, na.rm=T))
```

# assessment of univariate and multivariate normality with Mardia's test - work satisfation
```{r}
avg_only_sat <- sat_variables %>%
  select(swws_1:swws_5)

mvn(avg_only_sat)
```

Results indicate the violance of normality.


## 5. Final SEM MODEL
Using structural equation modelling for modeling the connection between 
burnout, work motivation and work satisfaction factors with satisfactory fit indices.

# Loading package
```{r}
library(lavaan)
```

# Final model
```{r}
my_model2 <- ' 
  # measurement model
burnout =~ Physical_B + Emotional_B + Cognitive_B
Physical_B =~ burn_1 + burn_3 + burn_5 + burn_8 + burn_10 + burn_12
Cognitive_B =~ burn_2 + burn_6 + burn_9 + burn_11 + burn_13
Emotional_B =~ burn_4 + burn_7 + burn_14

Intrin =~ mwms_3 + mwms_9 + mwms_15
Aut_Ext =~ mwms_5 + mwms_11 + mwms_17 + mwms_19
Amotiv =~ mwms_4 + mwms_10 + mwms_16
Cont_Ext =~ mwms_1 + mwms_2 + mwms_7 + mwms_8
Ext_Mat =~ mwms_6 + mwms_12 + mwms_18

work_satisf =~ swws_1 + swws_2 + swws_3 + swws_4 + swws_5
            swws_2 ~~ swws_5

# regressions
burnout ~ Amotiv + Ext_Mat + Cont_Ext + Aut_Ext + Intrin
work_satisf ~ Amotiv + Ext_Mat + Cont_Ext + Aut_Ext + Intrin
burnout ~~ work_satisf'
```


# model fit and summary of model fit
```{r}
my_model_fit2 <- cfa(my_model2, data = work_data, estimator = 'MLR')
summary(my_model_fit2, fit.measures=TRUE, standardized = TRUE)
```

Fit indices highlighted that the hypothesized representation of the constructs is satisfactory.


# Plotting the models with semPlot
```{r}
library(semPlot)
```

# Seeing the clean model structure
```{r}
semPaths(my_model_fit2)
```

# Getting standardized estimates
```{r}
semPaths(my_model_fit2, "std")
```


## 6. Results of final SEM Model

In the model the fit indices are acceptable (CFI = .909, TLI = .900, RMSEA = .054 [90% CI = .051-.057])

Both amotivation (beta = .27, p = .001) and controlled extrinsic motivation (beta = .215, p = .007) were 
related positively but weakly to burnout. Intrinsic motivation (beta = -.317, p < .001) has negative relation to
burnout. However, neither external material based motivaton (beta = -.012, p = .855), 
nor autonomous extrinsic motivation (beta = -.119, p = .143) related significantly to burnout.

Intrinsic motivation is positively related to work satisfaction (beta = .529, p < .001), while
amotivation has negative relation to that (beta = -.279, p < .001). 
External material based motivaton (beta = .103, p = .068), controlled extrinsic motivation (beta = .000, p = .998),
and autonomous extrinsic motivation (beta = .012, p = .831) does not have any relation to work satisfaction.

